### 2008-12-20 talk/ecoop2009/benchmarks.tex
some improvements, notes
comment out a possible confusing example, and add an XXX
a introduction about jit backends and jit layering, fix an XXX
hide the python-like syntax for factorial and fibonacci, add a link to the sources

\section{Benchmarks}

@@

Moreover, for each benchmark we also show the time taken by running the
equivalent program written in C\#.  By comparing the results against C\#, we
can see how far we are from the supposed optimal performances.  \anto{I
  don't really like the last sentence, but right now I can't think of another
  way to phrase it.  Rewording welcome :-)}

@@

To benchmark arithmetic operations between integers, we wrote a simple program
that computes the factorial of a given number.  The algorithm is
straightforward, and the loop contains only three operations: one
multiplication, one subtraction, and one comparison to check if we have
finished the job:

\begin{lstlisting}
def main(n):
    result = 1
    while n > 1:
        result = result * n
        n = n - 1
    return n
\end{lstlisting}

When doing plain interpretation, we need to create and destroy three temporary
objects at each iterations.  By contrast, the code generated by the JIT does
much better.  At the first iteration, the classes of the two operands of the
multiplication are promoted; then, the JIT compiler knows that both are
integers, so it can inline the code to compute the result.  Moreover, it can
\emph{virtualize} (see section \ref{sec:virtuals} all the temporary objects, because they never escape from
the inner loop.  The same remarks apply to the other two operations inside
the loop.

@@

Similarly, we wrote a program to calculate the $n_{th}$ Fibonacci number, for
which we can do the same reasoning as above:

\begin{lstlisting}
def main(n):
    a = 0
    b = 1
    while n > 1:
        sum = a + b
        a = b
        b = sum
        n = n - 1
    return b
\end{lstlisting}

@@

@@

Tables \ref{tab:factorial} and \ref{tab:fibo} show the time spent to calculate
the factorial and Fibonacci for various $n$.  As we can see, for small values
of $n$ the time spent running the JIT compiler is much higher than the time
spent to simply interpret the program.  This is an expected result, as till
now we only focused on optimizing the compiled code, not the compilation
process itself.

>>>>

\section{Benchmarks}
\label{sec:benchmarks}

@@

Moreover, for each benchmark we also show the time taken by running the
equivalent program written in C\# \footnote{The sources for both TLC and C\#
  programs are available at:
  http://codespeak.net/svn/pypy/extradoc/talk/ecoop2009/benchmarks/}

@@

To benchmark arithmetic operations between integers, we wrote a simple program
that computes the factorial of a given number.  The algorithm is
straightforward, thus we are not showing the source code.  The loop contains only three operations: one
multiplication, one subtraction, and one comparison to check if we have
finished the job.

\commentout{
\cfbolz{I think we can kill this for space reasons}
\anto{If we decide to keep them, we must remember to explain the python-like
  syntax, as it is no longer in tlc.tex}

\begin{lstlisting}
def main(n):
    result = 1
    while n > 1:
        result = result * n
        n = n - 1
    return n
\end{lstlisting}
}

When doing plain interpretation, we need to create and destroy three temporary
objects at each iteration.  By contrast, the code generated by the JIT does
much better.  At the first iteration, the classes of the two operands of the
multiplication are promoted; then, the JIT compiler knows that both are
integers, so it can inline the code to compute the result.  Moreover, it can
\emph{virtualize} (see section \ref{sec:virtuals} all the temporary objects, because they never escape from
the inner loop.  The same remarks apply to the other two operations inside
the loop.

@@

Similarly, we wrote a program to calculate the $n_{th}$ Fibonacci number, for
which we can do the same reasoning as above.

\commentout{
\cfbolz{I think we can kill this for space reasons}
\begin{lstlisting}
def main(n):
    a = 0
    b = 1
    while n > 1:
        sum = a + b
        a = b
        b = sum
        n = n - 1
    return b
\end{lstlisting}
}

@@

\cfbolz{can we make one figure of all the tables?}
\begin{table}[ht]
  \begin{tabular}{|l|r|r|r|r||r|r|}
    \hline
    \textbf{n} & 
    \textbf{Interp} &
    \textbf{JIT} &
    \textbf{JIT 2} &
    \textbf{C\#} &
    \textbf{Interp/JIT 2} &
    \textbf{JIT 2/C\#} \\
    \hline

@@

Tables \ref{tab:factorial} and \ref{tab:fibo} show the seconds spent to calculate
the factorial and Fibonacci for various $n$.  As we can see, for small values
of $n$ the time spent running the JIT compiler is much higher than the time
spent to simply interpret the program.  This is an expected result, as till
now we only focused on optimizing the compiled code, not the compilation
process itself.
